An Axiomatic Basis for Computer Programming: Review

### 2017-02-15 hoare/paper.tex
Write on tablet (about quadratic equation example and other)

For next disscussion suppose that we define right P and R from some
requirements. \\

@@

Then we could find values by some program Q.
And finally check whether equation holds or not in R: 
$$ R \equiv (X \subseteq \mathbb{R}) \wedge (\forall x \in X \bullet ax^2 + bx + c
= 0)$$

@@

In the article Hoare states four basic rules:
\begin{itemize}
\item
  Axiom of Assignment
\item
  Rule of Consequence
\item
  Rule of Composition
\item
  Rule of Iteration
\end{itemize}

Also it requires no side effects for Q (nothing changes except state which
we define in pre and postconditions). Because in other case, we couldn't
consider correctness of overall state. \\

@@

If we think about it in more details, we could decompose it to two sequence of
statements (rule of composition D2), in which first part hold some postcondition
R1 and second have R1 as precondition.

So nothing scary in that case if we still could prove correctness of our program.

>>>>

Or in other words, it helps us check that program do things right, but doesn't
approve that what we want, is right things.

For next discussion suppose that we define right P and R from some
requirements. \\

@@

Then we could find values by some program Q. Suppose as a result program returns
set of values X (Don't think about implementation yet).
And finally check whether equation holds or not in R: 
$$ R \equiv (X \subseteq \mathbb{R}) \wedge (\forall x \in X \bullet ax^2 + bx + c
= 0)$$

@@

In the '69 article Hoare states four basic rules:
\begin{itemize}
\item
  Axiom of Assignment
\item
  Rule of Consequence
\item
  Rule of Composition
\item
  Rule of Iteration
\end{itemize}
But sometimes stated another two rules - skip axiom and condition rule. First
one could be derived from axiom of assignment, and second from rule of
iteration.\\

All this rules implicitly assumed the absence of side effects for Q (nothing
changes except state which we define in pre and postconditions). Because in
other case, we couldn't consider correctness of overall state. \\

@@

If we think about it in more details, we could decompose it to two sequence of
statements (rule of composition D2), in which first part hold some postcondition
R1 and second have R1 as precondition. But R1 could be arbitrary, for example
$\not$ P.

As we can see, there is nothing scary in that case if we still could prove
correctness of our program.

So now lets return to our example about real roots.

Suppose we choose general quadratic formula for solving equation:

$$ x_{1,2} = \frac{-b +- \square{b^2 - 4ac}}{2a}$$

We know that $b^2 - 4ac$ not negative, but we don't know weather a is equals to
zero. In this case equation became linear, but still have solutions.

Lets try use substitution of x by this to prove that postcondition holds.
